---
title: 'Homework #3'
author: "Brooke Acosta, Emmy Park, Leah Martins-Krasner"
date: "2022-10-12"
output:
  html_document: default
  pdf_document: default
code_folding: hide
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(doBy)
library(kableExtra)
library(tidyverse)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Load Data
```{r}
hh <- read.csv("1_Household_Public.csv")
per <- read.csv("2_Person_Public.csv")
veh <- read.csv("3_Vehicle_Public.csv")
trip <- read.csv("4_Trip_Public.csv")
```

## Question 1
### Draw a random household from the second ten households in your data. If your data frame is named dat, this command will do it for you: sort(unique(dat$HH_ID))[sample(11:20, 1)]
### Write a brief description of the household. Be sure to include: their income, race, composition, county of residence, and whether they own a car.

```{r}
sort(unique(hh$HH_ID))[sample(11:20, 1)]
#Random household = 101113

# We drew Household 101113. The household is from Burlington County and is made up of one 55 to 64 year old male living in a single family, detached house. Their race was collected with the Pilot Record and is labelled as Non-Hispanic. They make between $75,000 and $99,000.  They own 1 car and 1 bike. 
```

## Question 2
### Describe the daily activities and travel of the household’s members using the trip data.
```{r}
subset(trip[c(2,3,15,18,28:35)], HH_ID == 101113)

# The single householder in Household 101113 took three trips on the study day: 

    # Trip 1 was done by car and had 3 stops. It started and ended at home with stops for a social visit and recreation (watch/observe movies, concert, sports         event, etc).  

    # Trip 2 was done by car and had 1 stop for a social visit. It lasted 1 hr 30 mins.  

    # Trip 3 was done by car and had 1 stop for online shopping for products, services, or goods.  
```

## Question 3
### Provide an estimate of the total number of bicycle trips represented by the survey data. Note that you should use the Person Weight to make this estimate. This number represents the total number of people each person in the survey is supposed to represent. For example, if a person’s Person Weight is 65, then that person represents 65 people in the Philadelphia region.

```{r}

trip <- merge(trip, hh[c("HH_ID", "H_COUNTY", "INCOME")], by = "HH_ID")

sum(trip$P_WEIGHT[trip$MODE_AGG=='2'],na.rm=T)
```

## Question 4 
### Make a table that shows the mode choice for residents from the county of the household from question 1. Note that you will have to make a choice about how to group the modes together. Note also that I asked for the home county not the origin of a trip county. Look for this in the data dictionary. 
```{r}
table(trip$MODE_AGG[hh$H_COUNTY == 34005])/sum(table(trip$MODE_AGG[hh$H_COUNTY == 34005]))*100

```

## Question 5 
### Make a graphic that shows the relationship between household income and the age of a household’s vehicle for all households. (Hint: merge the household data to the vehicle data with the merge() command. Use an internet search or type ?merge into the console to learn how to use the command.)

```{r}

hhveh <- merge(veh, hh[c("HH_ID", "H_COUNTY", "INCOME")], by = "HH_ID")

hhveh_fix <- filter(hhveh, INCOME < "98", YEAR < "5000")

library(ggplot2)

ggplot(hhveh_fix)+
  geom_point(aes(x = YEAR, 
                 y = INCOME))+
  geom_abline(intercept = 0, slope = 1)+
  labs(
    title = "Household Vehicle Age as a Function of Household Income",
    subtitle = "",
    caption = "",
    x="HH Vehicle Age", 
    y="HH Income")

```

## Question 6 
### Make a table or graphic that shows the relationship between household income and mode choice for all households. (The boxplot() command is a nice option, but a table is just fine.) Again, you must choose how to combine modes.

```{r}

hhtrip <- merge(trip, hh[c("HH_ID", "H_COUNTY", "INCOME")], by = "HH_ID")
hhtrip_fix <- filter(hhtrip, INCOME.x <"98")
Q6Table <- table(hhtrip_fix$INCOME.x,hhtrip_fix$MODE_AGG)
colnames(Q6Table) = c("Walk", "Bike", "Private Vehicle", "Private Transit", "Public Transit", "School Bus", "Other")
rownames(Q6Table) = c("$0 to $9,999","$10,000 to $24,999","$25,000 to $34,999", "$35,000 to $49,999", "$50,000 to $74,999","$75,000 to $99,999","$100,000 to $149,999","$150,000 to $199,999 ","$200,000 to $249,999","$250,000 or more") 
show(Q6Table)
```

## Question 7 
### The last three questions reference the output of an OLS regression model predicting daily boardings at US light rail stations:
#### a) According to the regression model, each additional job within a half mile of a station correlates with how many more daily transit riders per day?	

```{r}
# Each additional job within a half mile of a station correlates with 0.179 increase in daily transit riders.
```

#### b) If there are an average of 1,793 boardings per station and 3,130 jobs around each station, what percentage increase in ridership does the model predict due to a doubling of the number of jobs around a station (hint: elasticity).

```{r}
# Increase In Jobs * Elasticity = Increase In Boarding 

# 3130 * 0.179 = 560.27 increase in boardings 

# 560.27 increase in boardings + 1,793 boardings = 2,353.27 Boardings 

# There is a 31% increase in boardings when jobs double 
```

#### c) Describe the relationship between AM peak service frequency and light rail boardings.

```{r}
# There is a positive, linear relationship between AM peak service frequency and light rail boardings. For every increase in AM peak service frequency, boardings increase by 105.82.  The standard error is 9.887, so this increase in boarders could range from 95.33 to 115.71. 
```

